---
title: "class08"
author: 'Vince (PID: A15422556)'
date: "2/10/2022"
output:
  pdf_document: default
  html_document: default
---


## Unsupervised Learning Analysis of Human Breast Cancer Cells


Read data from the University of Wisconsin Medical Center on breast cancer patients

```{r}
# read.csv("WisconsinCancer.csv")

# Save your input data file into your Project directory
fna.data <- "WisconsinCancer.csv"

# Complete the following code to input the data and store as wisc.df
wisc.df <- read.csv(fna.data, row.names=1)
```


Remove the first column

```{r}
# We can use -1 here to remove the first column
wisc.data <- wisc.df[,-1]

# Create diagnosis vector for later
diagnosis <- as.factor(wisc.df$diagnosis)
```


## Exploratory Data Analysis


> Q1. How many observations are in this dataset? 31 observations.

```{r}
ncol(wisc.df)
```


> Q2. How many of the observations have a malignant diagnosis? 212 observations.

```{r}
# Table counts each combination of factors
table(wisc.df$diagnosis)

```


> Q3. How many variables/features in the data are suffixed with _mean? 10 variables/features.

```{r}
length(grep("_mean", colnames(wisc.df)))
```





## Principal Component Analysis (PCA)


> Q4. From your results, what proportion of the original variance is captured by the first principal components (PC1)? 44.3% of the original variance is captured by PC1.

```{r}
wisc.pr <- prcomp(wisc.data, scale=TRUE)
summary(wisc.pr)
```


> Q5. How many principal components (PCs) are required to describe at least 70% of the original variance in the data? 3 PCs are required to describe at least 70% of the original variance.

```{r}
summary(wisc.pr)
```


> Q6. How many principal components (PCs) are required to describe at least 90% of the original variance in the data? 7 PCs are required to describe at least 90% of the original variance.

```{r}
summary(wisc.pr)
```

###################################################################################

Make the main result: "PCA plot" (a.k.a. "score plot", PC1 vs PC2 plot).

```{r}
# PCA plot shows variation using PC1 and PC2, $x calls values from the data frame
plot(wisc.pr$x[,1:2], col=diagnosis)
```

###################################################################################

> Q7. What stands out to you about this plot? Is it easy or difficult to understand? Why? Plot is difficult to understand because everything is crowded together.

```{r}
biplot(wisc.pr)
```


> Q8. Generate a similar plot for principal components 1 and 3. What do you notice about these plots? There is less variance since PC3 does not contain as much variance as PC2 and there is clearer separation in the PC1 vs. PC2 plot.

```{r}
plot(wisc.pr$x[,1:2], col=diagnosis, xlab="PC1", ylab="PC2")
plot(wisc.pr$x[,c(1,3)], col=diagnosis, xlab="PC1", ylab="PC3")
```


ggplot

```{r}
# Create a data.frame for ggplot
df <- as.data.frame(wisc.pr$x)
df$diagnosis <- diagnosis

# Load the ggplot2 package
library(ggplot2)

# Make a scatter plot colored by diagnosis
ggplot(df) + 
  aes(PC1, PC2, col=diagnosis) + 
  geom_point()
```


Scree Plot

```{r}
# Calculate variance of each component
pr.var <- wisc.pr$sdev^2
head(pr.var)

# Variance explained by each principal component: pve
pve <- pr.var / sum(pr.var)

# Plot variance explained for each principal component
plot(pve, xlab = "Principal Component", 
     ylab = "Proportion of Variance Explained", 
     ylim = c(0, 1), type = "o")

# Alternative scree plot of the same data, note data driven y-axis
barplot(pve, ylab = "Percent of Variance Explained",
     names.arg=paste0("PC",1:length(pve)), las=2, axes = FALSE)
axis(2, at=pve, labels=round(pve,2)*100 )
```


factoextra Package

```{r}
## ggplot based graph
#install.packages("factoextra")
library(factoextra)
fviz_eig(wisc.pr, addlabels = TRUE)
```


> Q9. For the first principal component, what is the component of the loading vector (i.e. wisc.pr$rotation[,1]) for the feature concave.points_mean? The component of the loading vector for concave.points_mean is -0.26085376.

```{r}
wisc.pr$rotation[,1]
```


> Q10. What is the minimum number of principal components required to explain 80% of the variance of the data? 5 PCs are required to explain 80% of the variance of the data.

```{r}
summary(wisc.pr)
```





## Hierarchical Clustering


> Q11. Using the plot() and abline() functions, what is the height at which the clustering model has 4 clusters? At height 19 there are 4 clusters.

```{r}
# Scale the wisc.data data using the "scale()" function
data.scaled <- scale(wisc.data)
data.dist <- dist(data.scaled)
wisc.hclust <- hclust(data.dist, method="complete")

plot(wisc.hclust)
abline(h=19, col="red", lty=2)
```


> Q12. Can you find a better cluster vs diagnoses match by cutting into a different number of clusters between 2 and 10? Yes, cutting into 6 clusters reduces the ratio of benign vs malignant diagnoses where each cluster clearly favors one over the other. This improves the cluster vs diagnoses match.

```{r}
wisc.hclust.clusters <- cutree(wisc.hclust, k=6)
table(wisc.hclust.clusters, diagnosis)
```


> Q13. Which method gives your favorite results for the same data.dist dataset? Explain your reasoning. The "ward.D2" method is my favorite because I like how it minimizes variance and displays the data in a streamlined way.

```{r}
data.scaled <- scale(wisc.data)
data.dist <- dist(data.scaled)
wisc.hclust <- hclust(data.dist, method="ward.D2")
plot(wisc.hclust)
```





## K-means Clustering


> Q14. How well does k-means separate the two diagnoses? How does it compare to your hclust results? K-means separates the two diagnnoses fairly well and the results are similar to hclust results.

```{r}
wisc.km <- kmeans(scale(wisc.data), centers=2, nstart=20)
table(wisc.km$cluster, diagnosis)

table(wisc.hclust.clusters, wisc.km$cluster)
```





## Combining Methods


```{r}
wisc.pr.hclust <- hclust(dist(wisc.pr$x[,1:7]), method="ward.D2")
plot(wisc.pr.hclust)

grps <- cutree(wisc.pr.hclust, k=2)
table(grps)

table(grps, diagnosis)
plot(wisc.pr$x[,1:2], col=grps)

plot(wisc.pr$x[,1:2], col=diagnosis)

g <- as.factor(grps)
levels(g)
g <- relevel(g,2)
levels(g)
plot(wisc.pr$x[,1:2], col=g)
```


> Q15. How well does the newly created model with two clusters separate out the two diagnoses? It separates out the two clusters well.

```{r}
wisc.pr.hclust <- hclust(dist(wisc.pr$x[,1:7]), method="ward.D2")
wisc.pr.hclust.clusters <- cutree(wisc.pr.hclust, k=2)

table(wisc.pr.hclust.clusters, diagnosis)
```


> Q16. How well do the k-means and hierarchical clustering models you created in previous sections (i.e. before PCA) do in terms of separating the diagnoses? They do well in separating the diagnoses as the clusters show somewhat clear levels of either benign or malignant diagnoses.

```{r}
table(wisc.km$cluster, diagnosis)
table(wisc.hclust.clusters, diagnosis)
```





## Sensitivity/Specificity


> Q17. Which of your analysis procedures resulted in a clustering model with the best specificity? How about sensitivity? The best specificity resulted from hierarchical clustering, while the best specificity resulted from k-means clustering.





## Prediction


> Q18. Which of these new patients should we prioritize for follow up based on your results? We should prioritize following up with patients that make up cluster 2 (red, malignant).

```{r}
url <- "https://tinyurl.com/new-samples-CSV"
new <- read.csv(url)
npc <- predict(wisc.pr, newdata=new)
npc

plot(wisc.pr$x[,1:2], col=g)
points(npc[,1], npc[,2], col="blue", pch=16, cex=3)
text(npc[,1], npc[,2], c(1,2), col="white")
```
