---
title: "class14"
author: 'Vince (PID: A15422556)'
date: "3/3/2022"
output: 
  pdf_document:
    latex_engine: xelatex
---

Getting Started

> Q1. What column details the total number of people fully vaccinated? persons_fully_vaccinated

> Q2. What column details the Zip code tabulation area? zip_code_tabulation_area

> Q3. What is the earliest date in this dataset? 2021-01-05 

> Q4. What is the latest date in this dataset? 2022-03-01

```{r}
# Import vaccination data
vax <- read.csv("covid19vaccinesbyzipcode_test.csv")
head(vax)

head(vax$as_of_date)
tail(vax$as_of_date)
vax$as_of_date[nrow(vax)]
```


Skim Package

```{r}
library(skimr)
skimr::skim(vax)
```

> Q5. How many numeric columns are in this dataset? 9 columns

> Q6. Note that there are “missing values” in the dataset. How many NA values there in the persons_fully_vaccinated column? 18338

> Q7. What percent of persons_fully_vaccinated values are missing (to 2 significant figures)? 17.04%

```{r}
(sum(is.na(vax$persons_fully_vaccinated)) / 107604) * 100
```


> Q8. [Optional]: Why might this data be missing? It's possible the patients didn't follow up for their second dose so they aren't considered "fully vaccinated."


Working with Dates

> Q9. How many days have passed since the last update of the dataset? 2 days

> Q10. How many unique dates are in the dataset (i.e. how many different dates are detailed)? 61 unique entries

```{r}
library(lubridate)

vax$as_of_date <- ymd(vax$as_of_date)
today() - vax$as_of_date[nrow(vax)]

length(unique(vax$as_of_date))
```


Working with ZIP codes

```{r loading packages}
library(zipcodeR)
library(dplyr)
library(ggplot2)
```


Focus on the San Diego area.

```{r}
sd <- filter(vax, county == "San Diego")
dim(sd)
```


> Q11. How many distinct zip codes are listed for San Diego County? 107

```{r}
length(unique(sd$zip_code_tabulation_area))
```


> Q12. What San Diego County Zip code area has the largest 12 + Population in this dataset? 92154

```{r}
sd[which.max(sd$age12_plus_population), ]
```


> Q13. What is the overall average “Percent of Population Fully Vaccinated” value for all San Diego “County” as of “2022-03-01”? 70.5%

```{r}
#Filter to the day
sd.latest <- filter(sd, as_of_date=="2022-03-01")
mean(sd.latest$percent_of_population_fully_vaccinated, na.rm=TRUE)
```


> Q14. Using either ggplot or base R graphics make a summary figure that shows the distribution of Percent of Population Fully Vaccinated values as of “2022-03-01”?

```{r}
hist(sd.latest$percent_of_population_fully_vaccinated)
```


Focus on UCSD/La Jolla

```{r}
ucsd <- filter(sd, zip_code_tabulation_area=="92037")
ucsd[1,]$age5_plus_population
```


> Q15. Using ggplot make a graph of the vaccination rate time course for the 92037 ZIP code area:

```{r}
baseplot <- ggplot(ucsd) +
  aes(as_of_date,
      percent_of_population_fully_vaccinated) +
  geom_point() +
  geom_line(group=1) +
  ylim(c(0,1)) +
  labs(x="Date", y="Percent Vaccinated")

baseplot
```


> Q16. Calculate the mean “Percent of Population Fully Vaccinated” for ZIP code areas with a population as large as 92037 (La Jolla) as_of_date “2022-03-01”. Add this as a straight horizontal line to your plot from above with the geom_hline() function?

```{r}
vax.36 <- filter(vax, age5_plus_population > 36144 &
                as_of_date == "2022-03-01")

head(vax.36)
ave.36 <- mean(vax.36$percent_of_population_fully_vaccinated)

baseplot +
  geom_hline(yintercept = ave.36)
```


> Q17. What is the 6 number summary (Min, 1st Qu., Median, Mean, 3rd Qu., and Max) of the “Percent of Population Fully Vaccinated” values for ZIP code areas with a population as large as 92037 (La Jolla) as_of_date “2022-03-01”?

```{r}
summary(vax.36)
```


> Q18. Using ggplot generate a histogram of this data. 

```{r}
ggplot(vax.36) +
  aes(percent_of_population_fully_vaccinated) +
  geom_histogram() +
  xlim(c(0,1))
```


> Q19. Is the 92109 and 92040 ZIP code areas above or below the average value you calculated for all these above? The 92109 and 92040 zip code areas are both below the calculated average value.

```{r}
vax %>% filter(as_of_date == "2022-03-01") %>%  
  filter(zip_code_tabulation_area=="92109") %>%
  select(percent_of_population_fully_vaccinated)

vax %>% filter(as_of_date == "2022-03-01") %>%  
  filter(zip_code_tabulation_area=="92040") %>%
  select(percent_of_population_fully_vaccinated)
```


> Q20. Finally make a time course plot of vaccination progress for all areas in the full dataset with a age5_plus_population > 36144.

```{r}
vax.36.all <- filter(vax, age5_plus_population > 36144)


ggplot(vax.36.all) +
  aes(as_of_date,
      percent_of_population_fully_vaccinated, 
      group=zip_code_tabulation_area) +
  geom_line(alpha=0.2, color="blue") +
  ylim(c(0,1)) +
  labs(x="Date", y="Percent Vaccinated",
       title="Vaccination Rate Across California",
       subtitle="Only areas with a population above 36k are shown.") +
  geom_hline(yintercept = ave.36, linetype=2)
```


> Q21. How do you feel about traveling for Spring Break and meeting for in-person class afterwards? It might be okay, but it is still safer to stay at home.




