---
title: 'Class 6: R Functions'
author: 'Vince (PID: A15422556)'
date: "2/3/2022"
output:
  pdf_document: default
  html_document: default
---

Today we will explore R functions.

We will start with calculating a grade for these example students.
```{r}
# Example input vectors to start with 
student1 <-c(100, 100, 100, 100, 100, 100, 100, 90) 
student2 <-c(100, NA, 90, 90, 90, 90, 97, 80) 
student3 <-c(90, NA, NA, NA, NA, NA, NA, NA)
```

Use mean() to calculate an average.
```{r}
mean(student1, na.rm = TRUE)
```

Try is.na() on student 2.
```{r}
student2
is.na(student2)
```

Use result to get NA values (i.e. the TRUE positions).
```{r}
student2[is.na(student2)]
```

Put mean() and is.na() together.
```{r}
student3[is.na(student3)] <- 0
mean(student3)
```

Determine the lowest score.
```{r}
which.min(student1)
```

Remove the lowest score.
```{r}
student1[-which.min(student1)]
```

Calculate grade with the lowest score dropped.
```{r}
x <- student2
# Set NA values to zero
x[is.na(x)] <- 0
# Remove lowest score and calculate average
x <- x[-which.min(x)]
mean(x)
```

R Function
All R functions have 3 things:
-name (grade)
-input arguments (student scores)
-body (does the work)

```{r}
grade <- function(x) {
  x[is.na(x)] <- 0
  mean(x[-which.min(x)])
}
```



> Q1.

```{r}
gradebook <- read.csv("https://tinyurl.com/gradeinput", row.names=1)
head(gradebook)
```

Use the `apply()` function to grade the whole class.

> Q2.

```{r}
scores <- apply(gradebook, 1, grade)
which.max(scores)
```

> Q3.

```{r}
hardHw <- apply(gradebook, 2, median, na.rm = TRUE)
which.min(hardHw)

boxplot(gradebook)
```

> Q4.

```{r}
mask <- gradebook
mask[is.na(mask)] <- 0
mask
```

```{r}
apply(mask, 2, cor, scores)
```

> Q5. Knit
