---
title: "class10"
author: 'Vince (PID: A15422556)'
date: "2/17/2022"
output:
  pdf_document: default
  html_document: default
---

MXL Population

```{r}
mxl <- read.csv("373531-SampleGenotypes-Homo_sapiens_Variation_Sample_rs8067378.csv")

table(mxl$Genotype..forward.strand.) / nrow(mxl)
```


GBR Population

```{r}
gbr <- read.csv("373522-SampleGenotypes-Homo_sapiens_Variation_Sample_rs8067378.csv")

table(gbr$Genotype..forward.strand.) / nrow(gbr)
```


## Expression by Genotype Analysis

> Q13. Read file into R, determine sample size and median expression levels. Sample size for A/A is 108, for A/G is 233, and for G/G is 121. Median expression levels for A/A is 31.25, for A/G is 25.065, and for G/G is 20.074.

```{r}
x <- read.table("rs8067378_ENSG00000172057.6.txt")

table(x$geno)

print("G/G")
summary(x[x$geno == "G/G", 3])
print("A/G")
summary(x[x$geno == "A/G", 3])
print("A/A")
summary(x[x$geno == "A/A", 3])
```


> Q14. Is the mean expression different based on genotype? Yes, the SNP seems to cause a decrease in expression of ORMDL3.

## Summary Overview Figure

```{r}
library(ggplot2)

ggplot(x, aes(geno, exp, fill=geno)) + geom_boxplot(notch=TRUE)
```

